Passenger Prediction in Shared Accounts for Flight Service Recommendation

### 2016-07-17 Paper.tex
Add files via upload

\documentclass{llncs}
\bibliographystyle{splncs}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=blue,citecolor=blue]{hyperref}
\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{document}

@@

\author{Yafeng Zhao, Jian Cao}
\institute{Department of Computer Science and Engineering,\\
Shanghai Jiao Tong University, Shanghai 200240, China\\
\email{zyfgs2012@163.com, cao-jian@sjtu.edu.cn}}

@@

\textbf{Paper organization.} The rest of paper is structured as follows. We introduce related work in Section \ref{sec:relw}. In Section \ref{sec:mod}, we define the problem and describe the model. Besides, we also discuss how to do parameter inference. In Section \ref{sec:rec}, we propose a preference-based flight ticket recommendation approach, which also plays a role as a baseline method at evaluation stage. In Section \ref{sec:exp}, we introduce the experiment dataset, evaluation metrics, the experimental results of passenger prediction and flight recommendation. Finally, we conclude the paper in Section \ref{sec:con}.

@@

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items and it represents a latent feature. An account is a multinomial distribution over topics and it can represent the preferences of an account. Each purchase can be regarded a sample that selects a topic $z$ from account $a$, and takes an item from topic $z$.\par
When the topic model is applied to flight recommendation, there are some challenges. Firstly, the concepts behind flight tickets are quite difficult to be modeled. Conventional items only have static features, which means that the contents of an item are less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. Since the price factor plays a significant role in users' choice, so we can't consider the tickets with different prices as different items. Secondly, the amount of tickets for a flight is limited, which leads to a very sparse user-item matrix so that we can't use a collaborative-filtering like method directly. Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. The number of alternatives for each factor is also quite limited.\par
Our notations are summarized in table \ref{tab:not}.\par
\begin{table}[!htbp]
\centering
\caption{\label{tab:not}Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within the dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

@@

With the above generative model, $P(x_{oi}=p|\mathbf{p}_o)$ is assumed to be a uniform distribution over passenger list $P$. Each topic is drawn independently conditioned on $\Theta$ and $p_o$, and each word is drawn independently  conditioned on $\Phi$ and $z$. Equation \ref{eq:1} can be applied as the likelihood of all orders in a single account. If we treat $\Theta$ and $\Phi$ as random variables, our target is to estimate the Maximum A Posteriori for the generative model.\par

@@

Equation \ref{eq:2} represents the probability of assigning topic $t$ and passenger $p$ for $i$-th word in order $o$. $C^{WT}$ is the word-topic matrix, and $C_{wt}$ is the times that word $w$ is assigned to topic $t$ except for the current word. $C^{TP}$ represents the topic-passenger matrix, and $C_{tp}$ indicates the times that passenger $p$ is assigned to topic $t$ except for the current word $w_{oi}$. Moreover, $W$ is the size of vocabulary, $T$ represents the number of topics and $P$ is the number of passengers. From the sampling metric, we can estimate the topic-word distribution and passenger-topic distribution:\\

@@

The algorithm performs in three steps, i.e., initialization, sampling and updating respectively. At the first step, we assign each word in the corpus with random passengers and topics. For every sampling operation, a word in corpus is chosen, the probability of topic distribution and passenger distribution is calculated conditioned on the rest words in corpus by applying Equation \ref{eq:2}, with these two probabilities, we can sample a new topic and passenger for the current word. After several batch iterations, the passenger-topic matrix $\Theta$ and topic-word matrix $\Phi$ can be updated by applying Equation \ref{eq:3}, \ref{eq:4}. So the computational efficiency is the number of words multiplies number of topics, passengers and iteration times.\par

@@

In conclusion, the passenger prediction task can be partitioned into two steps. For the first step, a set of decisive factors are extracted based on domain  knowledge to generate a vocabulary. And the parameter $\Theta$ and $\Phi$ are trained through several batches of Gibbs sampling. In the second step, passengers of an anonymous order can be predicted by applying Equation \ref{eq:5}. Algorithm \ref{alg:1} describes the predicting process mathematically.\par

\begin{algorithm}[htb]
\caption{passengerPrediction}
\begin{algorithmic}[1]
\label{alg:1}
\REQUIRE
Account history order list $O$. \par
Predefined discrete factor list $F$ \par
A test anonymous order $o$
\ENSURE 
User's prediction list $Pl$
\STATE List $Pl \leftarrow \emptyset$;
\STATE stat word vector $\mathbf{W}$ for all orders;
\STATE Model $M \leftarrow trainATM(W)$;
\STATE calculate user's probability by Equation \ref{eq:5};
\FORALL{user : P}
\IF{$user.probability > \frac{1}{|P|}$}
\STATE $Pl.append(user,probability)$;
\ENDIF
\ENDFOR
\STATE normalize $Pl$;
\RETURN $Pl$;
\end{algorithmic} 
\end{algorithm}

\section{The Preference-based recommendation approach}
\label{sec:rec}
There is little existing work aiming at providing flight recommendation. In this section, we propose a user preference-based recommendation approach. In Section \ref{sec:exp}, we evaluate this method against some ranking strategies, such as price rank etc, to demonstrate that this method has a better accuracy. Then, we will apply the passenger prediction results to improve recommendation accuracy further.

@@

\section{Experiments}
\label{sec:exp}
In the following subsections, we introduce datasets and evaluation metrics. Then we evaluate the performance of passenger prediction and passenger prediction based recommendation respectively.
\subsection{Dataset}
In our experiments, we use a real flight ticket dataset, consisting of all submitted orders for two years. An order contains account id and all passengers' id. We use desensitized passenger id to distinguish individuals from each other. In our experiment, we select active users whose amount of historical orders reaches a threshold, thus passengers' preferences can be extracted with more confidence. In addition, many researches\cite{amy:guess}\cite{yutaka:modeling} on shared account recommendation generate an artificial dataset by composing single accounts' data. We also generate an artificial dataset combining single passengers' data from two different accounts. The overviews of these two datasets and the meta data of orders are summarized in table \ref{tab:dat} and table \ref{tab:mdat}.
\begin{table}[!htbp]
\centering
\caption{\label{tab:dat}Datasets}
\begin{tabular}{|c|c|c|c|} \hline
& \# accounts & \# passengers & \# orders\\ \hline
Real & 4632&7034&38907 \\ \hline
Artificial &1604&3208&29759\\ \hline
\end{tabular}
\end{table}

@@

\begin{equation}
Acc(u_i) =
\begin{cases}
1 & \mbox{if N =1}\\
1 - \frac{index-1}{N-1} & \mbox{if N >1}
\end{cases}
\end{equation}

@@

where $|M|$ is the number of total test orders, $Acc(u_i)$ represents the accuracy of recommendation based on the rank percentages of test orders. The metric MAP is the average of each user's accuracy, and is a measurement for general performance. The top-$N$ hit rate is the actual strategy for recommendation, we can stick up top-$N$ candidates based on the ranked list. For each purchase, if top-$N$ candidates hit the chosen order, we consider this recommendation has accuracy 1, else the accuracy is 0. We also calculate MAP for all accounts \par
For passenger prediction, we propose the following metric to evaluate the accuracy:\\
\begin{equation}
\label{eq:10}
Acc(u_i) = \frac{P' \cap A}{\sqrt{|P'| \times |A|}}
\end{equation}\par
Since we treat every account as a corpus, so a probabilistic distribution list is generated for all passengers contained in the account. At the first stage, we need to decide the passenger number for each test order. Assuming that there are total $|P|$ passengers in an account. So in average case, the probability of every passenger is $P_A = \frac{1}{|P|}$. We decide the passenger number by omitting passengers whose probability is not greater than $P_A$. We denote predicted passenger list as $P'$, actual passenger list as $A$. Thus we can get the prediction accuracy by Equation \ref{eq:10}. The numerator is the number of hit passengers, the denominator is penalty item for the size of predicted passenger list and predicted passenger list.

\subsection{Passenger Prediction}
\label{sub:pred}
We train a probabilistic model for every account where topic number selected varying ${5,10,\dots,50}$. We pick a test order for each account as a new document, then predict every passenger's probability of placing this order by Equation \ref{eq:5}. According to above evaluating metrics, the mean prediction accuracy rate of (a) real dataset and (b) artificial dataset is shown in Figure \ref{fig:pred}. The left sub-figure is the evaluation result on the real dataset and the right one is on the artificial dataset. The figure's horizontal axis shows the number of trained topics and the vertical axis shows the prediction accuracy computed by applying above mentioned metrics. The Random method means assigning random probabilities to passengers, we also record $10$ batch of random results as our approach does. For the artificial dataset, there are two passengers in every account and one passenger for every test order, so the accuracy of prediction is either 1.0 or 0. The results demonstrate that our model has a higher prediction precision than the random method on both datasets. Besides, when the topic number is 10 or 15, both datasets get the best performance while the performance goes down with the topic number increasing, which demonstrates that the size of corpus' vocabulary and number of words in each order also impact the optimal topic number. In the next subsection, we evaluate the recommendation performance with topic number fixed at 10.\par

>>>>

\documentclass{llncs}
\bibliographystyle{splncs03}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=blue,citecolor=blue]{hyperref}
\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{document}

@@

\author{Yafeng Zhao, Jian Cao\thanks{Corresponding Author}, Yudong Tan}
\institute{Department of Computer Science and Engineering,\\
Shanghai Jiao Tong University, Shanghai 200240, China and\\
Air Ticketing B.U., Ctrip.com International, Ltd.\\
\email{zyfgs2012@163.com,cao-jian@sjtu.edu.cn,ydtan@Ctrip.com}}

@@

\textbf{Paper organization.} The rest of paper is structured as follows. We introduce related work in Sect.\ref{sec:relw}. In Sect.\ref{sec:mod}, we define the problem and describe the model. Besides, we also discuss how to do parameter inference. In Sect.\ref{sec:rec}, we propose a preference-based flight ticket recommendation approach, which also plays a role as a baseline method at evaluation stage. In Sect.\ref{sec:exp}, we introduce the experiment dataset, evaluation metrics, the experimental results of passenger prediction and flight recommendation. Finally, we conclude the paper in Sect.\ref{sec:con}.

@@

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items and it represents a latent feature. An account is a multinomial distribution over topics and it can represent the preferences of an account. Each purchase can be regarded a sample that selects a topic $z$ from account $a$, and takes an item from topic $z$.\par
When the topic model is applied to flight recommendation, there are some challenges. Firstly, the concepts behind flight tickets are quite difficult to be modeled. Conventional items only have static features, which means that the contents of an item are less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. Since the price factor plays a significant role in users' choice, we can't consider the tickets with different prices as different items. Secondly, the amount of tickets for a flight is limited, which leads to a very sparse user-item matrix so that we can't use a collaborative-filtering like method directly. Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. The number of alternatives for each factor is also quite limited.\par
Our notations are summarized in Table \ref{tab:not}.\par
\begin{table}[!htbp]
\centering
\caption{\label{tab:not}Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within the dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

@@

With the above generative model, $P(x_{oi}=p|\mathbf{p}_o)$ is assumed to be a uniform distribution over passenger list $P$. Each topic is drawn independently conditioned on $\Theta$ and $p_o$, and each word is drawn independently  conditioned on $\Phi$ and $z$. Eq.\ref{eq:1} can be applied as the likelihood of all orders in a single account. If we treat $\Theta$ and $\Phi$ as random variables, our target is to estimate the Maximum A Posteriori for the generative model.\par

@@

Eq.\ref{eq:2} represents the probability of assigning topic $t$ and passenger $p$ for $i$-th word in order $o$. $C^{WT}$ is the word-topic matrix, and $C_{wt}$ is the times that word $w$ is assigned to topic $t$ except for the current word. $C^{TP}$ represents the topic-passenger matrix, and $C_{tp}$ indicates the times that passenger $p$ is assigned to topic $t$ except for the current word $w_{oi}$. Moreover, $W$ is the size of vocabulary, $T$ represents the number of topics and $P$ is the number of passengers. From the sampling metric, we can estimate the topic-word distribution and passenger-topic distribution:\\

@@

The algorithm performs in three steps, i.e., initialization, sampling and updating respectively. At the first step, we assign each word in the corpus with random passengers and topics. For every sampling operation, a word in corpus is chosen, the probability of topic distribution and passenger distribution is calculated conditioned on the rest words in corpus by applying Eq.\ref{eq:2}, with these two probabilities, we can sample a new topic and passenger for the current word. After several batch iterations, the passenger-topic matrix $\Theta$ and topic-word matrix $\Phi$ can be updated by applying Eq.\ref{eq:3}, Eq.\ref{eq:4}. So the computational efficiency is the number of words multiplies number of topics, passengers and iteration times.\par

@@

In conclusion, the passenger prediction task can be partitioned into two steps. For the first step, a set of decisive factors are extracted based on domain  knowledge to generate a vocabulary. And the parameter $\Theta$ and $\Phi$ are trained through several batches of Gibbs sampling. In the second step, passengers of an anonymous order can be predicted by applying Eq.\ref{eq:5}. Algorithm \ref{alg:1} describes the predicting process mathematically.\par

\begin{algorithm}[htb]
\caption{passengerPrediction}
\begin{algorithmic}[1]
\label{alg:1}
\REQUIRE
Account history order list $O$. \par
Predefined discrete factor list $F$ \par
A test anonymous order $o$
\ENSURE 
User's prediction list $Pl$
\STATE List $Pl \leftarrow \emptyset$;
\STATE stat word vector $\mathbf{W}$ for all orders;
\STATE Model $M \leftarrow trainATM(W)$;
\STATE calculate user's probability by Eq.\ref{eq:5};
\FORALL{user : P}
\IF{$user.probability > \frac{1}{|P|}$}
\STATE $Pl.append(user,probability)$;
\ENDIF
\ENDFOR
\STATE normalize $Pl$;
\RETURN $Pl$;
\end{algorithmic} 
\end{algorithm}

\section{The Preference-based recommendation approach}
\label{sec:rec}
There is little existing work aiming at providing flight recommendation. In this section, we propose a user preference-based recommendation approach. In Sect.\ref{sec:exp}, we evaluate this method against some ranking strategies, such as price rank etc, to demonstrate that this method has a better accuracy. Then, we will apply the passenger prediction results to improve recommendation accuracy further.

@@

\section{Experiments}
\label{sec:exp}
In the following subsections, we introduce datasets and evaluation metrics. Then we evaluate the performance of passenger prediction and passenger prediction based recommendation respectively.
\subsection{Dataset}
In our experiments, we use a real flight ticket dataset, consisting of all submitted orders for two years. An order contains account id and all passengers' id. We use desensitized passenger id to distinguish individuals from each other. In our experiment, we select active users whose amount of historical orders reaches a threshold, thus passengers' preferences can be extracted with more confidence. In addition, many researches\cite{amy:guess}\cite{yutaka:modeling} on shared account recommendation generate an artificial dataset by composing single accounts' data. We also generate an artificial dataset combining single passengers' data from two different accounts. The overviews of these two datasets and the meta data of orders are summarized in Table \ref{tab:dat} and Table \ref{tab:mdat}.
\begin{table}[!htbp]
\centering
\caption{\label{tab:dat}Datasets}
\begin{tabular}{|c|c|c|c|} \hline
& \# accounts & \# passengers & \# orders\\ \hline
Real & 4632&7034&38907 \\ \hline
Artificial &1604&3208&29759\\ \hline
\end{tabular}
\end{table}

@@

\begin{equation}
Acc(u_i) =
\begin{cases}
1 & \mbox{if N = 1}\\
1 - \frac{index-1}{N-1} & \mbox{if N $>$ 1}
\end{cases}
\end{equation}

@@

where $|M|$ is the number of total test orders, $Acc(u_i)$ represents the accuracy of recommendation based on the rank percentages of test orders. The metric MAP is the average of each user's accuracy, and is a measurement for general performance. The top-$N$ hit rate is the actual strategy for recommendation, we can stick up top-$N$ candidates based on the ranked list. For each purchase, if top-$N$ candidates hit the chosen order, we consider this recommendation has accuracy 1, else the accuracy is 0. We also calculate MAP for all accounts \par
For passenger prediction, we propose the following metric to evaluate the accuracy:\\
\begin{equation}
\label{eq:10}
Acc(u_i) = \frac{P' \cap A}{\sqrt{|P'| \times |A|}}
\end{equation}\par
Since we treat every account as a corpus, a probabilistic distribution list is generated for all passengers contained in the account. At the first stage, we need to decide the passenger number for each test order. Assuming that there are total $|P|$ passengers in an account. So in average case, the probability of every passenger is $P_A = \frac{1}{|P|}$. We decide the passenger number by omitting passengers whose probability is not greater than $P_A$. We denote predicted passenger list as $P'$, actual passenger list as $A$. Thus we can get the prediction accuracy by Eq.\ref{eq:10}. The numerator is the number of hit passengers, the denominator is penalty item for the size of predicted passenger list and predicted passenger list.

\subsection{Passenger Prediction}
\label{sub:pred}
We train a probabilistic model for every account where topic number selected varying ${5,10,\dots,50}$. We pick a test order for each account as a new document, then predict every passenger's probability of placing this order by Eq.\ref{eq:5}. According to above evaluating metrics, the mean prediction accuracy rate of (a) real dataset and (b) artificial dataset is shown in Fig.\ref{fig:pred}. The left sub-figure is the evaluation result on the real dataset and the right one is on the artificial dataset. The figure's horizontal axis shows the number of trained topics and the vertical axis shows the prediction accuracy computed by applying above mentioned metrics. The Random method means assigning random probabilities to passengers, we also record $10$ batch of random results as our approach does. For the artificial dataset, there are two passengers in every account and one passenger for every test order, so the accuracy of prediction is either 1.0 or 0. The results demonstrate that our model has a higher prediction precision than the random method on both datasets. Besides, when the topic number is 10 or 15, both datasets get the best performance while the performance goes down with the topic number increasing, which demonstrates that the size of corpus' vocabulary and number of words in each order also impact the optimal topic number. We will evaluate the recommendation performance with topic number fixed at 10.\par
